Cross-language Phonetic Transfer in Bilingual Speech

### 2020-09-11 paper_draft/main.tex
Added author info

\documentclass[11pt]{article}
\usepackage{apacite}
\usepackage{syntonly}
%\syntaxonly
%
%Margin - 1 inch on all sides
%
\usepackage[letterpaper]{geometry}
\usepackage{times}
\geometry{top=1.0in, bottom=1.0in, left=1.0in, right=1.0in}

@@

@@

\bibliographystyle{apacite}

@@

\item Long-term influence due to acquisition of multiple languages-- affects not only L2, but also L1. Bidirectional influence-- shapes phonetic categories in bilinguals, which differ from that of monolingual speakers\cite{caramazza1973acquisition, flege1987production, guion2003vowel}.

\item Temporary phonetic drift due to language exposure \cite{sancier1997gestural,chang2012rapid,tobin2017phonetic}
\end{enumerate}
These studies- look at the two languages of a bilingual speaker as used separately.
\citeA{grosjean1994going} first studied dynamic phonetic influence due to simultaneous activation of both languages of a bilingual speaker. Although they did not find any evidence of cross-language influence, and concluded that the switch between language systems is immediate and complete at the phonetic level, subsequent studies using similar paradigms have reported different results.

\paragraph{}These compare bilingual speakers’ productions during mixed-language usage to their own productions in a single language. The findings vary greatly—studies have shown unidirectional influence of L1 on L2 \cite{balukas2015spanish,antoniou2011inter,vsimavckova2015immediate,goldrick2014language}, L2 on L1 \cite{tsui2019impact,elias2017effects}, bidirectional convergence \cite{bullock2009trying, olson2016role}, divergence \cite{bullock2009trying,vsimavckova2018patterns} and no influence \cite{muldner2019phonetics,schwartz2015language}. Partly attributed to differences in methodology, populations, and language pairs, along with variation in individual switching strategies.

\paragraph{}Overall, the results suggest that cross-language phonetic influence in bilinguals increases during mixed-language processing, and both L1 and L2 sounds can show gradient phonetic shift in production as a result. The extent of shift is mediated by language-specific (sound system), context-specific (eg. language mode), and speaker-specific (proficiency, dominance) constraints. Studies eliciting spontaneous bilingual speech have also noted the influence of social factors \cite{khattab2009phonetic}.

\paragraph{}Most studies of this phenomenon have focused on consonant sounds, using differences in typical VOT of stop consonants to measure phonetic shift. Studies on vowel sounds are relatively fewer and much more recent, but suggest that vowels too show dynamic cross-language transfer. The studies so far report varied results: \citeA{elias2017effects,simonet2014phonetic,simonet2020increased} have reported L2 influence on L1 in code-switching (CS) tasks, whereas \citeA{muldner2019phonetics} found no significant increase in cross-language influence. As far as we are aware, there is no study yet on L1 influence on L2 vowels during mixed language use.

\paragraph{}The majority of studies reviewed so far have used CS (both spontaneous and scripted) to induce co-activation of languages. Other methods used include delayed repetition \cite{simonet2014phonetic}, reading single- or mixed-language word lists \cite{simonet2020increased}, and interpreting across languages \cite{vsimavckova2015immediate,vsimavckova2018patterns}. \citeA{olson2013bilingual} argues— CS (switching languages within a discourse, in connected speech) allows for planning and pragmatic considerations, which independently affect production, and might therefore obscure the underlying phonetic interaction due to mixed-language processing. They propose cued picture-naming as an alternative paradigm to induce switches in language—it lacks discourse context and predictability, so offers a more controlled way to observe phonetic effects due to language-processing manipulations. Following this, other studies \cite{goldrick2014language,tsui2019impact} have used this paradigm to study dynamic phonetic transfer in bilinguals.

\paragraph{}While phonetic transfer is usually measured at the point of the language-switch, previous  research has noted that the observed transfer could potentially have been triggered by two different sources— local phonetic processing (the switch itself), or global language co-activation (language mode \cite{grosjean1998studying} or context of the task). To examine this, several studies have manipulated language mode /context in addition to language-switching. Excepting one study (\citeNP{olson2013bilingual}; language context operationalized as proportion of words from each language during a test block) which found the balanced language context to mitigate transfer, results generally suggest that cross-language transfer due to switching either increases in a mixed language mode \cite{simonet2014phonetic}, or remains unaffected by it \cite{schwartz2015language,vsimavckova2015immediate,olson2016role,simonet2020increased}.

@@

\paragraph{}The results of \citeA{simonet2014phonetic} and \citeA{elias2017effects} suggest that phonetic interaction during switching affects vowel categories which are common across the two languages, as well as those which are not. For the purpose of this study, two English vowels were chosen:
(i) \wt{\ae}--exists as a category in both Bengali and General Indian English (GIE)\cite{masica1972sound}, and (ii) \wt{2}-- exists as a category in English, but not Bengali. It is assumed that the lack of a mid-central vowel in Bengali allows greater room for variablity due to shift (if any) in \nt{2} compared to \nt{\ae}. This criteria has been used to explain differences in the extent and direction of cross-language transfer in numerous studies measuring VOT. Therefore, we ask whether a parallel difference in the extent of shift is found in vowels.

@@

>>>>

\documentclass[11pt]{article}
% \usepackage{apacite}
\usepackage{syntonly}
%\syntaxonly
%
%Margin - 1 inch on all sides
%
\usepackage[letterpaper]{geometry}
\usepackage{times}
\geometry{top=1.0in, bottom=1.0in, left=1.0in, right=1.0in}

@@

\author{Auromita Mitra, Indranil Dutta}

@@

% \bibliographystyle{apacite}

@@

{}\item Long-term influence due to acquisition of multiple languages-- affects not only L2, but also L1. Bidirectional influence-- shapes phonetic categories in bilinguals, which differ from that of monolingual speakers\cite{caramazza1973acquisition, flege1987production, guion2003vowel}.

\item Temporary phonetic drift due to language exposure \cite{sancier1997gestural,chang2012rapid,tobin2017phonetic}
\end{enumerate}
These studies- look at the two languages of a bilingual speaker as used separately.
\cite{grosjean1994going} first studied dynamic phonetic influence due to simultaneous activation of both languages of a bilingual speaker. Although they did not find any evidence of cross-language influence, and concluded that the switch between language systems is immediate and complete at the phonetic level, subsequent studies using similar paradigms have reported different results.

These compare bilingual speakers’ productions during mixed-language usage to their own productions in a single language. The findings vary greatly—studies have shown unidirectional influence of L1 on L2 \cite{balukas2015spanish,antoniou2011inter,vsimavckova2015immediate,goldrick2014language}, L2 on L1 \cite{tsui2019impact,elias2017effects}, bidirectional convergence \cite{bullock2009trying, olson2016role}, divergence \cite{bullock2009trying,vsimavckova2018patterns} and no influence \cite{muldner2019phonetics,schwartz2015language}. Partly attributed to differences in methodology, populations, and language pairs, along with variation in individual switching strategies.

Overall, the results suggest that cross-language phonetic influence in bilinguals increases during mixed-language processing, and both L1 and L2 sounds can show gradient phonetic shift in production as a result. The extent of shift is mediated by language-specific (sound system), context-specific (eg. language mode), and speaker-specific (proficiency, dominance) constraints. Studies eliciting spontaneous bilingual speech have also noted the influence of social factors \cite{khattab2009phonetic}.

Most studies of this phenomenon have focused on consonant sounds, using differences in typical VOT of stop consonants to measure phonetic shift. Studies on vowel sounds are relatively fewer and much more recent, but suggest that vowels too show dynamic cross-language transfer. The studies so far report varied results: \cite{elias2017effects,simonet2014phonetic,simonet2020increased} have reported L2 influence on L1 in code-switching (CS) tasks, whereas \cite{muldner2019phonetics} found no significant increase in cross-language influence. As far as we are aware, there is no study yet on L1 influence on L2 vowels during mixed language use.

The majority of studies reviewed so far have used CS (both spontaneous and scripted) to induce co-activation of languages. Other methods used include delayed repetition \cite{simonet2014phonetic}, reading single- or mixed-language word lists \cite{simonet2020increased}, and interpreting across languages \cite{vsimavckova2015immediate,vsimavckova2018patterns}. \cite{olson2013bilingual} argues- CS (switching languages within a discourse, in connected speech) allows for planning and pragmatic considerations, which independently affect production, and might therefore obscure the underlying phonetic interaction due to mixed-language processing. They propose cued picture-naming as an alternative paradigm to induce switches in language-it lacks discourse context and predictability, so offers a more controlled way to observe phonetic effects due to language-processing manipulations. Following this, other studies \cite{goldrick2014language,tsui2019impact} have used this paradigm to study dynamic phonetic transfer in bilinguals.

While phonetic transfer is usually measured at the point of the language-switch, previous research has noted that the observed transfer could potentially have been triggered by two different sources— local phonetic processing (the switch itself), or global language co-activation (language mode \cite{grosjean1998studying} or context of the task). To examine this, several studies have manipulated language mode /context in addition to language-switching. Excepting one study (\cite{olson2013bilingual}; language context operationalized as proportion of words from each language during a test block) which found the balanced language context to mitigate transfer, results generally suggest that cross-language transfer due to switching either increases in a mixed language mode \cite{simonet2014phonetic}, or remains unaffected by it \cite{schwartz2015language,vsimavckova2015immediate,olson2016role,simonet2020increased}.

@@

\paragraph{}The results of \cite{simonet2014phonetic} and \cite{elias2017effects} suggest that phonetic interaction during switching affects vowel categories which are common across the two languages, as well as those which are not. For the purpose of this study, two English vowels were chosen:
(i) \wt{\ae}--exists as a category in both Bengali and General Indian English (GIE)\cite{masica1972sound}, and (ii) \wt{2}-- exists as a category in English, but not Bengali. It is assumed that the lack of a mid-central vowel in Bengali allows greater room for variablity due to shift (if any) in \nt{2} compared to \nt{\ae}. This criteria has been used to explain differences in the extent and direction of cross-language transfer in numerous studies measuring VOT. Therefore, we ask whether a parallel difference in the extent of shift is found in vowels.

@@

\bibliographystyle{unsrt}
